Instructions for TTO 2019 Proceedings

### 2021-07-05 article/draft.tex
update

One of Facebook’s measures to regulate misinformation is to reduce the spread of misleading content through a ranking system. Facebook ranks each post and/or ad by assigning to it a relevancy score, where a high score leads to a high likelihood of the post and/or the ad to appear on a user's newsfeed. doing so, Facebook can make a post or a whole account less visible by decreasing the relevancy score of its content; this is precisely the $reduce$ measure.\footnote{Lyons, T. (2018, May 22). The three-part recipe for cleaning up your news feed. Facebook Newsroom. \href{https://about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}{about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}.} Such {\it reduce} measure should be visible in the engagement metrics of a given post. That is, if a post reaches less users because it has a lower ranking, then it is less likely to receive likes, comments and shares, relative to a post with a higher ranking.

To illustrate, we investigate the case of the website $Infowars$. Owned by Alex Jones, Infowars is usually described as a misinformation website: it is on the Wikimedia global spam blacklist\footnote{See \href{https://meta.wikimedia.org/wiki/Spam\_blacklist}{meta.wikimedia.org/wiki/Spam\_blacklist}.}, was classified among websites that have posted deceptive content according to FactCheck.org\footnote{See \href{https://www.factcheck.org/2017/07/websites-post-fake-satirical-stories/}{https://www.factcheck.org/2017/07/websites-post-fake-satirical-stories/}.}, and its factual reporting has been rated {\it very low} by the Media Bias / Fact Check resource.\footnote{See the Iffy.news page: \href{https://mediabiasfactcheck.com/infowars-alex-jones/}{https://mediabiasfactcheck.com/infowars-alex-jones/}.} On May 2, 2019, Facebook announced they would prohibit users from sharing Infowars content unless they are explicitly condemning the material.\footnote{See \href{https://www.wired.com/story/facebook-bans-alex-jones-extremists/}{https://www.wired.com/story/facebook-bans-alex-jones-extremists/}.}

To verify the measure, we used the CrowdTangle API at the ‘/posts/search’ endpoint\footnote{see the documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Search}{github.com/CrowdTangle/API/wiki/Search}.} to collect the 37,242 Facebook public posts that shared a link containing ‘infowars.com’ and that were published between January 1, 2019 and December 31, 2020.\footnote{We found in the collected data some Facebook posts that did not directly share an Infowars link (but rather a YouTube or Facebook video containing an Infowars link in its description), thus we excluded such posts from our data to keep only the 27,721 posts directly sharing an Infowars link.} The command used can be found \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_infowars_data.sh}{here}.

The number of public posts sharing an Infowars link remained globally stable in 2019 (Figure \ref{infowars1} top panel), thus Facebook’s announced ‘ban’ did not appear to prevent users from sharing an Infowars link. Interestingly a clear drop in engagement was observed on May 2, 2019 (Figure \ref{infowars1} bottom panel). The number of reactions per post decreased by -94\% when we compared the two months before and after May 2, 2019, the shares decreased by -96\%, and the comments by -93\%.

@@

	\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/infowars/fb_infowars_1.png}
	\end{multicols}

	\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/infowars/fb_infowars_2.png} 
	\end{multicols}

	\caption{Public Facebook posts sharing an Infowars link in 2019 and 2020 gathered from the CrowdTangle API. The red line marks the date of May 2, 2019, when Facebook has announced the ban regarding Infowars. (Top panel) Number of daily posts. (Bottom panel) Engagement metrics (average number of reactions, shares and comments per post). }
	\label{infowars1}
\end{figure}

This would suggest that Facebook’s measure taken in May 2019 is actually not a ban, but would rather correspond to a ‘reduce’ action, as users could still post Infowars links, but these posts generated vastly less engagement. It should be noted that the engagement metrics increased again at the end of 2019 / beginning of 2020, suggesting that the reducing measure may have been lifted a few months after its implementation.

As CrowdTangle is tracking posts only from certain public groups and pages, we also used the ‘/search/articles’ endpoint of the Buzzsumo API, to gather more complete Facebook data. We collected the engagement data for the 14,232 articles crawled by Buzzsumo from the Infowars website between January 1, 2019 and Decembre 31, 2020. The command can be found here.
We observe that the articles published after May 2, 2019 received less Facebook engagement than the ones published before (Figure X5), with a percentage change of -97\% for the reactions, -59\% for the shares and -97\% for the comments. An increase in engagement was also observed in 2020. It reinforces the hypothesis that Facebook reduced the reach of posts sharing Infowars links during a few months in 2019.

We found irregularities in the number of Infowars articles collected from Buzzsumo. While Infowars usually publishes 20-30 articles per day, only 53 articles were collected in the 31-day period between June 11 to July 11, 2020. A temporary crawling problem coming from Buzzsumo may have caused this lack of data. Because no database is perfect, we would like to highlight the importance of cross-checking information between different sources when possible.

@@

\begin{figure}
	\centering
		\begin{multicols}{1}	
			\includegraphics[scale=0.3]{./img/globalresearch_2021-06-14.png}
		\end{multicols}

		\begin{multicols}{1}
			\includegraphics[scale=0.3]{./img/globalresearch_14_06_2021_16pm_UTC.png} 
		\end{multicols}
		\caption{Screenshots taken on June 14, 2021. }
		\label{fig3}
\end{figure}

@@

\begin{figure*}
\centering
	\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/globalresearch/sum_globalresearch.ca_6_months.jpg} 
		\includegraphics[scale=0.2]{./img/globalresearch/engagement_sum_rolling_1_globalresearch.ca.jpg}
	\end{multicols}

	\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/globalresearch/tweet.png} 
		\includegraphics[scale=0.2]{./img/globalresearch/warning.png}
	\end{multicols}
\caption{}
\label{fig4}
\end{figure*}

The website $globalresearch.ca$ is linked to the Twitter account {$@CRG\_CRM$}; that was recently suspended.\footnote{We noticed the message about the account suspension on May 25, 2021. But to the best of our knowledge, no official communication by Twitter has announced the suspension nor the exact date at which it was implemented. Hence the account may have gotten suspended anytime between April 15, 2021 and May 25, 2021 (see the suspension screenshot in panel $(a)$ of figure \ref{fig3}). Furthermore, $globalresearch.ca$ has multiple failed fact-checks, see \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/}} When a user searches via the twitter search-box a url link of this website, no results appear as shown in the screenshot in panel $(b)$ of figure \ref{fig3}, taken on June 14, 2021. To further investigate the possible implementation of a reduced visibility measure, we search via the Twitter API for tweets, excluding retweets, containing the query {\it globalresearch.ca} from January 1, 2021 until June 10, 2021.  As shown in panel (a) in figure \ref{fig4}, we find a strictly positive number of tweets containing the URL link {\it globalresearch.ca} throughout May 2021 and the first week of June 2021. Hence, the visibility of tweets containing this URL link has been reduced because users can no longer access tweets containing the URL link {\it globalresearch.ca} via the search box. Nevertheless users are not restrained from posting tweets containing this URL, as shown in the screenshot in panel $(c)$ of figure \ref{fig4}, found by taking the tweet ID of one of the collected tweets via the Twitter API. Furthermore, those collected Tweets have strictly positive engagement metrics as shown in panel $(b)$ of figure \ref{fig4}. Hence, the users who tweet articles from the {\it globalresearch.ca} website receive positive engagement from their own followers. Finally, when a user attempts to click on the URL link {\it globalresearch.ca} contained in the Tweet, a warning message appears and indicates that the link may be unsafe (see screenshot in panel (d) in figure \ref{fig4}).

@@

\end{itemize}
}

>>>>

One of Facebook’s measures to regulate misinformation is to reduce the spread of misleading content through their ranking system. Facebook ranks each post and/or ad by assigning to it a relevancy score, where a high score leads to a high likelihood of the post and/or the ad to appear on a user's newsfeed. Doing so, Facebook can make a post or a whole account less visible by decreasing the relevancy score of its content; this is precisely the $reduce$ measure.\footnote{Lyons, T. (2018, May 22). The three-part recipe for cleaning up your news feed. Facebook Newsroom. \href{https://about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}{about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}.} This measure can be verified by looking at the number of views (reach) of a post, but this metric is not available via the CrowdTangle API or on Buzzsumo. Hence we can indirectly investigate the $reduce$ measure by looking at the engagement metrics (likes, comments, shares) related to a given post; which are available via the CrowdTangle API and Buzzsumo. If a post reaches less users because it has a lower ranking, then it is less likely to receive likes, comments and shares, relative to a post with a higher ranking.

%phrase pas accès au reach, mais seulement les engagement metrics
% Owned by Alex Jones, Infowars is usually described as a misinformation website: it is on the Wikimedia global spam blacklist\footnote{See \href{https://meta.wikimedia.org/wiki/Spam\_blacklist}{meta.wikimedia.org/wiki/Spam\_blacklist}.},
%CrowdTangle and Buzzsumo provide access to the interaction metrics on content (engagement), but not to the number of views (or reach), therefore we can only verify the {\it reduce} measure indirectly through a decrease in engagement. If Facebook was reducing the reach of a content via a lower ranking, we would expect the content to be less likely to receive likes, comments and shares, relative to a post with a higher ranking.

To illustrate, we investigate the case of the website $Infowars$. This website appears in the Misinformation Directory of FactCheck.org, among other websites who have posted deceptive content\footnote{See \href{https://www.factcheck.org/2017/07/websites-post-fake-satirical-stories/}{https://www.factcheck.org/2017/07/websites-post-fake-satirical-stories/}.}. Furthermore, the factual reporting of $Infowars$ has been rated {\it very low} by the Media Bias / Fact Check resource of Iffy.news.\footnote{See the Iffy.news page: \href{https://mediabiasfactcheck.com/infowars-alex-jones/}{https://mediabiasfactcheck.com/infowars-alex-jones/}.}

On May 2, 2019, Facebook announced they would prohibit users from sharing Infowars content unless, they are explicitly condemning the material.\footnote{See \href{https://www.wired.com/story/facebook-bans-alex-jones-extremists/}{https://www.wired.com/story/facebook-bans-alex-jones-extremists/} and \href{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}{about.fb.com/news/2018/08/enforcing-our-community-standards/}.} To verify the measure, we used the ``/posts/search'' endpoint\footnote{see the documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Search}{github.com/CrowdTangle/API/wiki/Search}.} of the CrowdTangle API, to collect $37$ $242$ Facebook public posts that had shared a URL link containing ``infowars.com'', published between January $1$, $2019$ and December $31$, $2020$.\footnote{We found in the collected data some Facebook posts that did not directly share an Infowars link (but rather a YouTube or Facebook video containing an Infowars link in its description), thus we excluded such posts from our data to keep only the $27$ $721$ posts directly sharing an Infowars link.} The command used can be found in the following Github repository: \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_infowars_data.sh}{link}. The number of public posts sharing an $Infowars$ link remained globally stable throughout $2019$ (see figure \ref{infowars1} top panel). Thus the measure announced by Facebook doesn't seem to have prevented users from sharing an $Infowars$ link. Nevertheless, a clear drop in engagement was observed on May 2, 2019 (see figure \ref{infowars1} bottom panel). The number of reactions, shares and comments per post have decreased respectively  by $-94\%$,  $-96\%$ and $-93\%$  when comparing the two months before and after May 2, 2019. This suggests the measure taken by Facebook in May $2019$, is not a ban per se, but rather a $reduce$ measure. This is because users could still post $Infowars$ links, but these posts generated less engagement. It should be noted that the engagement metrics increased again by the end of $2019$ / beginning of $2020$, suggesting that the $reduce$ measure may have been lifted a few months after its implementation.

@@

	\begin{multicols}{1}
		\includegraphics[scale=0.35]{./img/infowars/fb_infowars_1.png}
	\end{multicols}

	\begin{multicols}{1}
		\includegraphics[scale=0.35]{./img/infowars/fb_infowars_2.png} 
	\end{multicols}

	\caption{Public Facebook posts sharing an Infowars link in $2019$ and $2020$ collected from the CrowdTangle API. The red line marks the date of May $2$, 2019, when Facebook has announced the ban regarding Infowars. (Top panel) Number of daily posts. (Bottom panel) Engagement metrics: average number of reactions, shares and comments per post. }
	\label{infowars1}
\end{figure}

As CrowdTangle is tracking posts only from certain public groups and pages, we also used the ``/search/articles'' endpoint of the Buzzsumo API, to gather a richer Facebook dataset. We collected the engagement data for the $14$ $232$ articles crawled by Buzzsumo from the $Infowars$ website between January $1$, $2019$ and December $31$, $2020$.  \footnote{The command can be found here: \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_infowars\_data.py}{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_infowars\_data.py}.} We observe that the articles published after May $2$, $2019$ received less Facebook engagement than the ones published before (see figure \ref{infowars2}), with a percentage change of -$97\%$ for the reactions, -$59\%$ for the shares and -$97\%$ for the comments. An increase in engagement was also observed in $2020$. It reinforces the hypothesis that Facebook reduced the reach of posts sharing Infowars links only during a few months in $2019$.

{\color{pink} Finally, we found irregularities in the number of Infowars articles collected from Buzzsumo. While Infowars usually publishes $20$ to $30$ articles per day, only $53$ articles were collected in the $31$-day period between June $11$ to July $11$, 2020. A temporary crawling problem coming from Buzzsumo may have caused this lack of data. Because no database is perfect, we would like to highlight the importance of cross-checking information between different sources when possible.}

	\begin{multicols}{1}
		\includegraphics[scale=0.35]{./img/infowars/fb_infowars_3.png}
	\end{multicols}

	\caption{Facebook engagement metrics (average number of reactions, shares and comments per article) for the Infowars articles published in $2019-2020$ and gathered from the Buzzsumo API. The red line marks the date of May $2$, $2019$, when Facebook has announced the ban regarding Infowars.}
	\label{infowars2}
\end{figure}

@@

\begin{figure}
	\centering
		\begin{multicols}{1}	
			\includegraphics[scale=0.35]{./img/globalresearch_2021-06-14.png}
		\end{multicols}

		\begin{multicols}{1}
			\includegraphics[scale=0.35]{./img/globalresearch_14_06_2021_16pm_UTC.png} 
		\end{multicols}
		\caption{Screenshots taken on June 14, 2021. Top panel: screenshot that shows the account of $globalresearch.ca$ suspended on Twitter. Bottom panel: screenshot that shows that no results can be found when searching for $globalresearch.ca$.    }
		\label{fig3}
\end{figure}

@@

\begin{figure*}
\centering
	\begin{multicols}{1}
		\includegraphics[scale=0.35]{./img/globalresearch/sum_globalresearch.ca_6_months.jpg} 
		\includegraphics[scale=0.35]{./img/globalresearch/engagement_sum_rolling_1_globalresearch.ca.jpg}
	\end{multicols}
\caption{Top panel: daily number of Tweets, excluding retweets, containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $10$, $2021$. Bottom panel: engagement metrics of Tweets containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $10$, $2021$. Data collected via the Twitter API v2 on June $16$, 2021.   }
\label{fig4}
\end{figure*}

The website $globalresearch.ca$ is linked to the Twitter account {$@CRG\_CRM$}; which was recently suspended.\footnote{We noticed the message about the account suspension on May 25, 2021. But to the best of our knowledge, no official communication by Twitter has announced the suspension nor the exact date at which it was implemented. Hence the account may have gotten suspended anytime between April 15, 2021 and May 25, 2021 (see the suspension screenshot in panel $(a)$ of figure \ref{fig3}). Furthermore, $globalresearch.ca$ has multiple failed fact-checks, see \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/}} When a user searches via the twitter search-box for any URL link of this website, no results appear as shown in the screenshot in panel $(b)$ of figure \ref{fig3}, taken on June $14$, $2021$. To further investigate the possible implementation of a reduced visibility measure, we search via the Twitter API for tweets, excluding retweets, containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $10$, $2021$.  As shown in panel (a) in figure \ref{fig4}, we find a strictly positive number of tweets containing the URL link {\it globalresearch.ca} throughout May $2021$ and the first week of June $2021$. Hence, the visibility of tweets containing this URL link has been reduced because users can no longer access tweets containing the URL link {\it globalresearch.ca} via the search box. Nevertheless users are not restrained from posting tweets containing this URL, as shown in the screenshot in panel $(a)$ of figure \ref{fig4bis}, found by taking the tweet ID of one of the collected tweets via the Twitter API. Furthermore, those collected Tweets have strictly positive engagement metrics as shown in panel $(b)$ of figure \ref{fig4}. Hence, the users who tweet articles from the {\it globalresearch.ca} website receive tweet level engagement from their own followers. Finally, when a user attempts to click on the URL link {\it globalresearch.ca} contained in the Tweet, a warning message appears and indicates that the link may be unsafe (see screenshot in panel (b) in figure \ref{fig4bis}).

\begin{figure*}
\centering	
	\begin{multicols}{1}
		\includegraphics[scale=0.35]{./img/globalresearch/tweet.png} 
		\includegraphics[scale=0.35]{./img/globalresearch/warning.png}
	\end{multicols}
\caption{}
\label{fig4bis}
\end{figure*}

@@

	\item Pas accès au ``reach" !  ranking not available. Fb nous a donné certaines données (les données Condor) et dedans on a le nombre de clicks par exemple

	\item ``All four Pages have been unpublished for repeated violations of Community Standards and accumulating too many strikes. While much of the discussion around Infowars has been related to false news, which is a serious issue that we are working to address by demoting links marked wrong by fact checkers and suggesting additional content, none of the violations that spurred today’s removals were related to this." \href{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}{Newsroom}.
\end{itemize}
}

### 2021-07-05 article/test.tex
update

%To illustrate we provide two examples linked to the Twitter accounts of the websites $globalresearch.ca$ and $off$-$guardian.org$. Both websites have several failed fact-checks according to \href{https://iffy.news}{iffy.news} - a website which provides a database of websites with low factual reporting levels.\footnote{For $globalresearch.ca$ see \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/} and for $off-guardian.org$ see \href{https://mediabiasfactcheck.com/offguardian/}{https://mediabiasfactcheck.com/offguardian/}.  }
\begin{figure*}
\centering
	\begin{multicols}{2}
		\includegraphics[scale=0.2]{./img/globalresearch/sum_globalresearch.ca_6_months.jpg} 
		\includegraphics[scale=0.2]{./img/globalresearch/engagement_sum_rolling_1_globalresearch.ca.jpg}
	\end{multicols}
	\begin{multicols}{2}
		\includegraphics[scale=0.2]{./img/globalresearch/tweet.png} 
		\includegraphics[scale=0.2]{./img/globalresearch/warning.png}
	\end{multicols}
%\includegraphics[scale=0.3]{./img/globalresearch/sum_rolling_7_globalresearch.jpg} 
%\includegraphics[scale=0.3]{./img/globalresearch/sum_rolling_7_globalresearchall_time.jpg}
\caption{}
\label{fig4}
\end{figure*}

>>>>

%To illustrate we provide two examples linked to the Twitter accounts of the websites $globalresearch.ca$ and $off$-$guardian.org$. Both websites have several failed fact-checks according to \href{https://iffy.news}{iffy.news} - a website which provides a database of websites with low factual reporting levels.\footnote{For $globalresearch.ca$ see \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/} and for $off-guardian.org$ see \href{https://mediabiasfactcheck.com/offguardian/}{https://mediabiasfactcheck.com/offguardian/}.  }
\begin{figure}
\centering
	\begin{multicols}{1}
		\includegraphics[scale=0.35]{./img/globalresearch/sum_globalresearch.ca_6_months.jpg} 
		\includegraphics[scale=0.35]{./img/globalresearch/engagement_sum_rolling_1_globalresearch.ca.jpg}
	\end{multicols}
\caption{}
\label{fig4}
\end{figure}

\begin{figure}
\centering	
	\begin{multicols}{1}
		\includegraphics[scale=0.35]{./img/globalresearch/tweet.png} 
		\includegraphics[scale=0.35]{./img/globalresearch/warning.png}
	\end{multicols}
\caption{}
\label{fig4bis}
\end{figure}
